### 2014-12-30 Cuatrimestres/1/Matematicas/capitulos/03-diferenciales.tex

%-------------------------------------------------------------------------------
%	EMPIEZA SECCION
%-------------------------------------------------------------------------------

\section{Solución de una ecuación diferencial}

%-------------------------------------------------------------------------------
%	EMPIEZA SECCION
%-------------------------------------------------------------------------------

\section{Existencia y unicidad de las soluciones de una ecuación diferencial}

%-------------------------------------------------------------------------------
%	EMPIEZA SECCION
%-------------------------------------------------------------------------------

\section{Soluciones aproximadas}

%-------------------------------------------------------------------------------
%	EMPIEZA SECCION
%-------------------------------------------------------------------------------

\section{Relación entre soluciones aproximadas y exactas}